### 2024-08-06 code/eda-code/eda2.qmd
commit changes

---
title: "Exploratory analysis script"
date: "2024-02-07"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata2.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary_df, file = summarytable_file)
```

```{r}
summary_data = summary(mydata)
print(summary_data)
# save to file
summarytable_file = here("results", "tables", "summarytable2.rds")
saveRDS(summary_data, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures.

# Data exploration through figures

Find outliers.

BMI first.

```{r}
p1 <- mydata %>% ggplot(aes(x = "", y = bmi)) + geom_boxplot() + labs(title = "BMI Outliers")
plot(p1)
figure_file = here("results", "figures","bmi_outliers.png")
ggsave(filename = figure_file, plot=p1) 
```

Now Charges.

```{r}
p2 <- mydata %>% ggplot(aes(x = "", y = charges)) + geom_boxplot() + labs(title = "Charges Outliers")
plot(p2)
figure_file = here("results", "figures","charges-outliers.png")
ggsave(filename = figure_file, plot=p2) 
```

Now histogram of insurance charges.

```{r}
p3 <- mydata %>% ggplot(aes(x=charges)) +
  geom_histogram(binwidth=2000, fill="blue", color="black", alpha=0.7) +
  labs(title="Distribution of Insurance Charges", x="Charges", y="Frequency") +
  theme_minimal()
plot(p3)
figure_file = here("results", "figures", "Charges-histogram.png")
ggsave(filename = figure_file, plot=p3) 
```

Now let's look at a boxplot for charges by smoking status

```{r}
p4 <- mydata %>% ggplot(aes(x=smoker, y=charges, fill=smoker)) +
  geom_boxplot() +
  labs(title="Insurance Charges by Smoking Status", x="Smoker", y="Charges") +
  scale_fill_manual(values=c("yes"="red", "no"="green")) +
  theme_minimal()
plot(p4)
figure_file = here("results", "figures", "charges-smoking-boxplot.png")
ggsave(filename = figure_file, plot=p4) 
```

Boxplot for charges by region

```{r}
p5 <- mydata %>% ggplot(aes(x=region, y=charges, fill=region)) +
  geom_boxplot() +
  labs(title="Insurance Charges by Region", x="Region", y="Charges") +
  theme_minimal()
plot(p5)
figure_file = here("results", "figures", "charges-region-boxplot.png")
ggsave(filename = figure_file, plot=p5) 
```

Scatter plot for charges by age
```{r}
p6 <- mydata %>% ggplot(aes(x=age, y=charges)) +
  geom_point(alpha=0.5, color="blue") +
  labs(title="Insurance Charges by Age", x="Age", y="Charges") +
  theme_minimal()
plot(p6)
figure_file = here("results", "figures", "charge-age-boxplot.png")
ggsave(filename = figure_file, plot=p6) 
```

Scatter plot for charges by BMI
```{r}
p7 <- mydata %>% ggplot(aes(x=bmi, y=charges)) +
  geom_point(alpha=0.5, color="green") +
  labs(title="Insurance Charges by BMI", x="BMI", y="Charges") +
  theme_minimal()
plot(p7)
figure_file = here("results", "figures", "charge-bmi-scatter.png")
ggsave(filename = figure_file, plot=p7) 
```

Bar Plot if average charges by number of children
```{r}
p8 <- mydata %>% ggplot(aes(x=as.factor(children), y=charges, fill=as.factor(children))) +
  stat_summary(fun.y=mean, geom="bar") +
  labs(title="Average Insurance Charges by Number of Children", x="Number of Children", y="Average Charges") +
  theme_minimal()
plot(p8)
figure_file = here("results", "figures", "children-bar.png")
ggsave(filename = figure_file, plot=p8) 
```